---
title: "Finding Optimal Testing Strategies"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Finding Optimal Testing Strategies}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 6,
  eval = FALSE  # Set to TRUE once data is integrated
)
```

## Introduction

This vignette demonstrates a complete workflow for systematically evaluating and identifying optimal testing strategies for quarantined individuals. The methods shown here are based on the analysis scripts in `inst/examples/` and illustrate how to:

1. Generate all possible testing schedules within constraints
2. Evaluate each schedule in parallel for computational efficiency
3. Identify optimal schedules for different scenarios and quarantine durations
4. Visualize and compare results

This approach allows policy-makers to balance test resources against transmission reduction.

## Setup

```{r setup}
library(isolatr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(foreach)
library(doParallel)
```

## Define Simulation Parameters

First, define the epidemiological and intervention parameters:

```{r parameters}
# Epidemiological parameters
n.ind <- 100000  # Large sample for stable estimates
TP <- 1.2        # Transmission potential
k <- 0.2         # Dispersion parameter (superspreading)

# Vaccination parameters (NSW example)
p.vac.idx <- 0.723   # Index case vaccination probability
p.vac.sc <- 0.723    # Secondary case vaccination probability
vacc.cor <- 0.9      # Household vaccination correlation

# Vaccine effectiveness
VE.trans <- 0.4      # Against transmission
VE.inf <- 0.72       # Against infection

# State
the.state <- "NSW"

# TTIQ scenarios to evaluate
the.scenarios <- c("optimal", "partial", "current_nsw_case_init")

# Quarantine durations to evaluate
quarantine.durations <- c(14, 7)
```

## Generate Testing Schedules

Use `test.designs()` to systematically generate all valid testing schedules:

```{r generate_schedules}
# Maximum possible test days
max_days <- max(quarantine.durations)

# Minimum spacing between tests (e.g., 2 days)
min.space <- 2

# Generate all schedules with 1, 2, or 3 tests
all.times3 <- test.designs(
  times = 1:max_days,
  n = 3,
  min.space = min.space,
  t1.prior.to = max_days
)

all.times2 <- test.designs(
  times = 1:max_days,
  n = 2,
  min.space = min.space,
  t1.prior.to = max_days
)

all.times1 <- test.designs(
  times = 1:max_days,
  n = 1,
  min.space = min.space,
  t1.prior.to = max_days
)

# Combine all schedules
all.times <- all.times3 %>%
  filter(Var1 <= 5) %>%
  add_row(
    all.times2 %>%
      mutate(Var3 = Inf) %>%
      filter(Var1 <= 5)
  ) %>%
  add_row(
    all.times1 %>%
      mutate(Var2 = Inf, Var3 = Inf)
  )

nn <- nrow(all.times)
print(paste("Evaluating", nn, "testing schedules"))
```

## Run Simulations for Each Scenario

Now we loop over quarantine durations and TTIQ scenarios, evaluating all testing schedules in parallel:

```{r run_simulations}
results_list <- list()

for (hq.d in quarantine.durations) {
  for (ts in the.scenarios) {

    cat(paste("\nEvaluating", ts, "scenario with", hq.d, "day quarantine\n"))

    # For 7-day quarantine, filter out schedules that extend beyond day 7
    if (hq.d == 7 && ncol(all.times) == 3) {
      all.times_filtered <- all.times %>%
        select(-Var3) %>%
        distinct() %>%
        filter(Var1 <= hq.d & (Var2 <= hq.d | is.infinite(Var2)))
      nn <- nrow(all.times_filtered)
    } else {
      all.times_filtered <- all.times
    }

    # Step 1: Run base simulation (once per scenario/duration combination)
    set.seed(1)  # For reproducibility
    CQ.output <- CQ.sim2(
      n.ind = n.ind,
      TP = TP,
      k = k,
      p.vac.idx = p.vac.idx,
      p.vac.sc = p.vac.sc,
      vacc.cor = vacc.cor,
      VE.trans = VE.trans,
      VE.inf = VE.inf,
      quarantine.duration = hq.d,
      the.scenario = ts,
      the.state = the.state
    )

    # Step 2: Evaluate all testing schedules in parallel
    cl <- makeCluster(4)  # Use 4 cores (adjust based on your machine)
    registerDoParallel(cl)

    tic <- Sys.time()

    output <- foreach(
      i = 1:nn,
      .combine = rbind,
      .packages = c("tidyverse", "isolatr")
    ) %dopar% {
      CQ.sim.test.times(
        CQ.sim.output = CQ.output,
        n.ind = n.ind,
        test.times = all.times_filtered[i, ] %>% as.numeric(),
        VE.trans = VE.trans,
        the.scenario = ts
      )
    }

    toc <- Sys.time() - tic
    cat(paste("Completed in", round(toc, 2), attr(toc, "units"), "\n"))

    stopCluster(cl)

    # Combine schedules with results
    results <- cbind(all.times_filtered, output) %>%
      mutate(
        scenario = ts,
        q.duration = hq.d
      )

    results_list[[paste(ts, hq.d, sep = "_")]] <- results
  }
}

# Combine all results
all_results <- bind_rows(results_list)
```

## Process and Analyze Results

```{r process_results}
# Clean up column names
colnames_base <- c("t1", "t2", "t3")
if (ncol(all.times) <= 3) {
  colnames(all_results)[1:3] <- colnames_base[1:3]
}

# Add number of tests
all_results <- all_results %>%
  mutate(
    t2 = case_when(
      is.na(t2) ~ Inf,
      TRUE ~ t2
    ),
    t3 = case_when(
      is.na(t3) ~ Inf,
      TRUE ~ t3
    )
  ) %>%
  mutate(n.tests = 3 - is.infinite(t2) - is.infinite(t3))

# Identify optimal schedules
optimal_schedules <- all_results %>%
  group_by(scenario, q.duration, n.tests) %>%
  filter(IPq == min(IPq)) %>%
  mutate(TT = paste0("(", t1, ",", t2, ",", t3, ")")) %>%
  mutate(TT = gsub(pattern = ",Inf", replacement = "", x = TT)) %>%
  select(scenario, q.duration, n.tests, TT, IPq, sdIPq, mean.cases)

print(optimal_schedules)
```

## Visualizations

### 1. Optimal Strategies Bar Plot

```{r bar_plot}
optimal_schedules %>%
  mutate(
    scenario = factor(
      scenario,
      levels = c("optimal", "current_nsw_case_init", "partial"),
      labels = c("Optimal delays", "NSW case-initiated delays", "Partial delays"),
      ordered = TRUE
    )
  ) %>%
  mutate(
    q.duration = factor(
      q.duration,
      levels = c("14", "7"),
      labels = c("14-days", "7-days"),
      ordered = TRUE
    )
  ) %>%
  ggplot() +
  aes(x = TT, y = IPq, fill = as.factor(n.tests)) +
  geom_col(position = position_dodge(preserve = "single")) +
  scale_x_discrete("Test Strategy") +
  scale_y_continuous("IPq (Infection Potential in Quarantine)") +
  scale_fill_discrete("Number of Tests") +
  facet_grid(q.duration ~ scenario, scales = "free_x") +
  theme_minimal() +
  theme(
    text = element_text(size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

### 2. Scatter Plot: Test Timing vs IPq

For each number of tests, visualize how test timing affects IPq:

```{r scatter_plots}
# Identify optimal schedules for highlighting
the.opt <- all_results %>%
  mutate(id = row_number()) %>%
  mutate(
    q.duration = factor(
      q.duration,
      levels = c("7", "14"),
      labels = c("7-day", "14-day"),
      ordered = TRUE
    )
  ) %>%
  group_by(q.duration, scenario, n.tests) %>%
  filter(IPq == min(IPq)) %>%
  pivot_longer(cols = t1:t3) %>%
  filter(is.finite(value))

# Plot for 3-test schedules
plot.the.scenario <- "optimal"

all_results %>%
  filter(scenario == plot.the.scenario, n.tests == 3) %>%
  mutate(id = row_number()) %>%
  mutate(
    q.duration = factor(
      q.duration,
      levels = c("7", "14"),
      labels = c("7-day", "14-day"),
      ordered = TRUE
    )
  ) %>%
  pivot_longer(cols = t1:t3) %>%
  ggplot() +
  aes(x = value, y = IPq, group = id) +
  geom_line(alpha = 0.1) +
  geom_point(aes(colour = name), alpha = 0.6) +
  geom_point(
    data = the.opt %>% filter(scenario == plot.the.scenario, n.tests == 3),
    aes(x = value, y = IPq, group = id),
    size = 3,
    colour = "black"
  ) +
  geom_line(
    data = the.opt %>% filter(scenario == plot.the.scenario, n.tests == 3),
    aes(x = value, y = IPq, group = id),
    size = 1,
    colour = "black"
  ) +
  scale_y_continuous("IPq") +
  scale_x_continuous("Test day for PCC", breaks = seq(1, 14, by = 1)) +
  facet_grid(q.duration ~ .) +
  scale_colour_manual(
    "Test Number",
    values = c("darkviolet", "green4", "darkorange"),
    labels = c("1", "2", "3")
  ) +
  theme_minimal() +
  theme(legend.position = "top")
```

### 3. Efficiency Analysis

Identify schedules that perform within 2% of optimal:

```{r efficiency}
near_optimal <- all_results %>%
  group_by(scenario, q.duration, n.tests) %>%
  arrange(IPq) %>%
  mutate(efficiency = IPq / min(IPq)) %>%
  filter(efficiency <= 1.02) %>%
  mutate(
    T1 = paste0(min(t1), "-", max(t1)),
    T2 = paste0(min(t2), "-", max(t2)),
    T3 = paste0(min(t3), "-", max(t3))
  ) %>%
  mutate(
    T2 = gsub("Inf-Inf", "-", T2),
    T3 = gsub("Inf-Inf", "-", T3)
  ) %>%
  select(scenario, q.duration, n.tests, T1, T2, T3, efficiency) %>%
  distinct() %>%
  arrange(desc(q.duration), desc(n.tests))

print(near_optimal)
```

This shows the range of test timings that achieve near-optimal performance, providing flexibility for implementation.

## Key Insights

From this systematic evaluation, we can derive several insights:

1. **Optimal test timing varies by scenario**: Faster TTIQ systems allow later first tests
2. **Test spacing matters**: Tests should be strategically spaced to capture infections at different stages
3. **Diminishing returns**: Adding a 3rd test provides smaller marginal benefit than adding a 2nd
4. **Robustness**: Multiple schedules often perform nearly as well as the single optimum

## Comparing Different Quarantine Policies

You can extend this framework to compare alternative policies, such as differential quarantine by vaccination status:

```{r differential_policy, eval=FALSE}
# Example: 7-day quarantine for vaccinated, 14-day for unvaccinated
# This requires running separate simulations for each group
# See inst/examples/runCq_7unv_14v_quarantine.R for full implementation
```

## Exporting Results

```{r export}
# Save optimal schedules
write.csv(
  optimal_schedules,
  "optimal_testing_schedules.csv",
  row.names = FALSE
)

# Save near-optimal ranges
write.csv(
  near_optimal,
  "near_optimal_testing_ranges.csv",
  row.names = FALSE
)
```

## Computational Considerations

For large-scale evaluations:

- **Parallelization** reduces runtime dramatically (4 cores ≈ 4x speedup)
- **Sample size** (n.ind): 100,000 provides stable estimates; 10,000 faster for exploration
- **Number of schedules**: Grows combinatorially with max days and tests
- **Memory**: Each simulation stores detailed secondary infection data

Typical runtime: ~30-60 minutes for complete analysis (all scenarios × durations) with 4 cores.

## Next Steps

- See `vignette("model-details")` for understanding how IPq is calculated
- See `vignette("quick-start-guide")` for basic usage
- Adapt this workflow for your own epidemic parameters and constraints

## References

This workflow is based on:

> [Citation to paper describing the optimization approach]

For parallel computing in R:
> `?foreach`, `?doParallel`
