---
title: "Quick Start Guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quick Start Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

```{r setup}
library(isolatr)
library(dplyr)
library(ggplot2)
```

## Introduction

The **isolatr** package provides tools for simulating COVID-19 transmission during quarantine and evaluating different testing strategies. This quick start guide will walk you through the basic workflow.

## Basic Workflow

The typical workflow consists of three steps:

1. **Simulate** transmission dynamics using `CQ.sim2()`
2. **Evaluate** testing strategies using `CQ.sim.test.times()` or `CQ.sim.notest()`
3. **Compare** results across different strategies

### Step 1: Run the Simulation

First, we simulate transmission from 1000 index cases (primary close contacts) under quarantine:

```{r simulate, eval=FALSE}
set.seed(42)  # For reproducibility

sim_results <- CQ.sim2(
  n.ind = 1000,                  # Number of index cases to simulate
  TP = 3.0,                      # Transmission potential (mean secondary cases)
  k = 0.25,                      # Dispersion (lower = more superspreading)
  p.vac.idx = 0.7,               # Probability index case is vaccinated
  p.vac.sc = 0.7,                # Probability secondary cases are vaccinated
  vacc.cor = 0.8,                # Household vaccination correlation
  VE.trans = 0.5,                # Vaccine effectiveness against transmission
  VE.inf = 0.7,                  # Vaccine effectiveness against infection
  quarantine.duration = 14,      # Quarantine period (days)
  the.scenario = "optimal",      # TTIQ scenario
  the.state = "NSW"              # Australian state
)
```

The output is a data frame with one row per secondary infection:

```{r examine, eval=FALSE}
head(sim_results)

# Check infection types
table(sim_results$who)
```

### Step 2: Evaluate a Testing Strategy

Now we can evaluate how effective a specific testing schedule would be:

```{r test_strategy, eval=FALSE}
# Test on days 1, 3, and 6
strategy_1 <- CQ.sim.test.times(
  CQ.sim.output = sim_results,
  n.ind = 1000,
  test.times = c(1, 3, 6),
  VE.trans = 0.5,
  the.scenario = "optimal"
)

print(strategy_1)
```

The output shows:
- **IPq**: Infection Potential in Quarantine (lower is better)
- **sdIPq**: Standard deviation
- **mean.cases**: Mean undetected cases per index case

### Step 3: Compare Multiple Strategies

Let's compare several testing strategies:

```{r compare, eval=FALSE}
# Define strategies
strategies <- list(
  "Days 1,3,6" = c(1, 3, 6),
  "Days 2,7" = c(2, 7),
  "Days 1,5,10" = c(1, 5, 10),
  "Day 6 only" = 6
)

# Evaluate each
results <- lapply(names(strategies), function(name) {
  out <- CQ.sim.test.times(
    CQ.sim.output = sim_results,
    n.ind = 1000,
    test.times = strategies[[name]],
    VE.trans = 0.5,
    the.scenario = "optimal"
  )
  out$strategy <- name
  out$n_tests <- length(strategies[[name]])
  return(out)
})

# Add no-testing baseline
baseline <- CQ.sim.notest(sim_results, n.ind = 1000,
                          VE.trans = 0.5, the.scenario = "optimal")
baseline$strategy <- "No testing"
baseline$n_tests <- 0

# Combine and display
comparison <- bind_rows(results, baseline)
print(comparison %>% arrange(IPq))
```

## Understanding the Parameters

### Epidemiological Parameters

- **TP** (Transmission Potential): Mean number of secondary cases per index case. Typical values: 1-5
- **k** (Dispersion): Controls overdispersion. Lower values (0.1-0.3) indicate more superspreading
- **VE.trans**: Vaccine effectiveness against transmission (0-1)
- **VE.inf**: Vaccine effectiveness against infection (0-1)

### TTIQ Scenarios

Three scenarios are available, representing different public health system performance:

- **optimal**: Fast contact tracing, rapid testing, minimal delays
- **partial**: Moderate delays in tracing and testing
- **current_nsw_case_init**: NSW baseline performance

Each scenario has different distributions for:
- Time to isolation (contact tracing speed)
- Test turnaround time
- Interview and notification delays

### State-Specific Parameters

The `the.state` parameter determines the household size distribution. Valid options:
- "NSW", "VIC", "QLD", "SA", "WA", "TAS", "NT", "ACT"
- Or full names: "New South Wales", "Victoria", etc.

## Interpreting Results

### Infection Potential in Quarantine (IPq)

IPq represents the expected number of tertiary infections from undetected secondary cases. It accounts for:
- How long infections remain undetected
- Generation interval distribution
- Transmission potential
- Vaccine effectiveness

**Lower IPq = more effective strategy**

### Mean Cases

The mean number of undetected secondary cases per index case. This complements IPq by showing the raw number of cases that escape detection.

## Next Steps

- See the **Model Details** vignette for in-depth explanation of the simulation model
- See the **Testing Strategies** vignette for systematic strategy comparison methods
- See the **Advanced Usage** vignette for customization and extensions

## Quick Reference

```{r reference, eval=FALSE}
# Basic simulation
sim <- CQ.sim2(n.ind, TP, k, p.vac.idx, p.vac.sc, vacc.cor,
               VE.trans, VE.inf, quarantine.duration,
               the.scenario, the.state)

# Evaluate testing
results <- CQ.sim.test.times(sim, n.ind, test.times, VE.trans, the.scenario)

# Evaluate no testing
baseline <- CQ.sim.notest(sim, n.ind, VE.trans, the.scenario)
```
