---
title: "Quick Start Guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quick Start Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

```{r setup}
library(isolatr)
library(dplyr)
library(ggplot2)
```

## Introduction

The **isolatr** package provides tools for simulating infectious disease transmission during isolation/quarantine and evaluating different testing strategies. While the default parameters are calibrated to COVID-19, the package can be used for any communicable disease by adjusting the epidemiological parameters.

This quick start guide walks you through the basic workflow using COVID-19 as an example.

## Basic Workflow

The typical workflow consists of three steps:

1. **Simulate** transmission dynamics using `CQ.sim2()`
2. **Evaluate** testing strategies using `CQ.sim.test.times()` or `CQ.sim.notest()`
3. **Compare** results across different strategies

### Step 1: Run the Simulation

First, we simulate transmission from 1000 index cases (primary close contacts) under isolation:

```{r simulate, eval=FALSE}
set.seed(42)  # For reproducibility

# COVID-19 example with default epidemiological parameters
sim_results <- CQ.sim2(
  n.ind = 1000,                  # Number of index cases to simulate
  TP = 3.0,                      # Transmission potential (mean secondary cases)
  k = 0.25,                      # Dispersion (lower = more superspreading)
  p.vac.idx = 0.7,               # Probability index case is vaccinated
  p.vac.sc = 0.7,                # Probability secondary cases are vaccinated
  vacc.cor = 0.8,                # Household vaccination correlation
  VE.trans = 0.5,                # Vaccine effectiveness against transmission
  VE.inf = 0.7,                  # Vaccine effectiveness against infection
  quarantine.duration = 14,      # Isolation period (days)
  the.scenario = "optimal",      # TTIQ scenario
  region = "NSW"                 # Region for household size distribution
)
```

The output is a data frame with one row per secondary infection:

```{r examine, eval=FALSE}
head(sim_results)

# Check infection types
table(sim_results$who)
```

### Step 2: Evaluate a Testing Strategy

Now we can evaluate how effective a specific testing schedule would be:

```{r test_strategy, eval=FALSE}
# Test on days 1, 3, and 6
strategy_1 <- CQ.sim.test.times(
  CQ.sim.output = sim_results,
  n.ind = 1000,
  test.times = c(1, 3, 6),
  TP = 3.0,
  VE.trans = 0.5,
  the.scenario = "optimal"
)

print(strategy_1)
```

The output shows:
- **IPq**: Infection Potential in Quarantine (lower is better)
- **sdIPq**: Standard deviation
- **mean.cases**: Mean undetected cases per index case

### Step 3: Compare Multiple Strategies

Let's compare several testing strategies:

```{r compare, eval=FALSE}
# Define strategies
strategies <- list(
  "Days 1,3,6" = c(1, 3, 6),
  "Days 2,7" = c(2, 7),
  "Days 1,5,10" = c(1, 5, 10),
  "Day 6 only" = 6
)

# Evaluate each
results <- lapply(names(strategies), function(name) {
  out <- CQ.sim.test.times(
    CQ.sim.output = sim_results,
    n.ind = 1000,
    test.times = strategies[[name]],
    TP = 3.0,
    VE.trans = 0.5,
    the.scenario = "optimal"
  )
  out$strategy <- name
  out$n_tests <- length(strategies[[name]])
  return(out)
})

# Add no-testing baseline
baseline <- CQ.sim.notest(sim_results, n.ind = 1000, TP = 3.0,
                          VE.trans = 0.5, the.scenario = "optimal")
baseline$strategy <- "No testing"
baseline$n_tests <- 0

# Combine and display
comparison <- bind_rows(results, baseline)
print(comparison %>% arrange(IPq))
```

## Using Custom Disease Parameters

The package supports modeling any infectious disease by customizing the epidemiological parameters. Here's an example using influenza-like parameters:

```{r influenza_example, eval=FALSE}
set.seed(42)

# Influenza example with custom parameters
flu_results <- CQ.sim2(
  n.ind = 1000,
  TP = 1.5,                      # Lower transmission potential
  k = 0.5,                       # Less overdispersion than COVID-19
  p.vac.idx = 0.4,
  p.vac.sc = 0.4,
  vacc.cor = 0.6,
  VE.trans = 0.3,                # Lower vaccine effectiveness
  VE.inf = 0.5,
  quarantine.duration = 7,       # Shorter isolation period

  # TTIQ scenario
 the.scenario = "optimal",

  # Custom household distribution (e.g., UK data)
  hh_probs = c(0.28, 0.35, 0.18, 0.12, 0.05, 0.015, 0.004, 0.001),

  # Influenza-specific distributions
  inc_meanlog = 0.34,            # ~1.4 day median incubation
  inc_sdlog = 0.42,
  gi_meanlog = 0.91,             # ~2.5 day median generation interval
  gi_sdlog = 0.52
)
```

### Common Disease Parameters

| Disease | Incubation (meanlog, sdlog) | Generation Interval (meanlog, sdlog) |
|---------|----------------------------|--------------------------------------|
| COVID-19 (default) | 1.63, 0.5 (~5.1 day median) | 1.376, 0.567 (~3.96 day median) |
| Influenza | 0.34, 0.42 (~1.4 day median) | 0.91, 0.52 (~2.5 day median) |
| SARS | 1.39, 0.51 (~4 day median) | 2.0, 0.45 (~7.4 day median) |

## Understanding the Parameters

### Epidemiological Parameters

- **TP** (Transmission Potential): Mean number of secondary cases per index case. Typical values: 1-5
- **k** (Dispersion): Controls overdispersion. Lower values (0.1-0.3) indicate more superspreading
- **VE.trans**: Vaccine/immunity effectiveness against transmission (0-1)
- **VE.inf**: Vaccine/immunity effectiveness against infection (0-1)
- **inc_meanlog, inc_sdlog**: Lognormal parameters for incubation period
- **gi_meanlog, gi_sdlog**: Lognormal parameters for generation interval

### TTIQ Scenarios

Three built-in scenarios represent different public health system performance:

- **optimal**: Fast contact tracing, rapid testing, minimal delays
- **partial**: Moderate delays in tracing and testing
- **baseline**: Baseline system performance (formerly "current_nsw_case_init")

Each scenario has different distributions for:
- Time to isolation (contact tracing speed)
- Test turnaround time
- Interview and notification delays

You can also create custom scenarios:

```{r custom_scenario, eval=FALSE}
# Create a custom scenario for a different context
custom_scenario <- create_scenario_config(
  name = "rapid_response",
  active_detection = list(type = "lognormal", meanlog = 0.8, sdlog = 0.4),
  passive_detection = list(type = "lognormal", meanlog = 1.2, sdlog = 0.5),
  test_turnaround = list(type = "exponential", rate = 2),
  interview_delay = list(type = "exponential", rate = 1),
  notification_delay = list(type = "exponential", rate = 0.8)
)
```

### Household Size Distribution

The `region` parameter uses built-in Australian state data, or you can provide custom probabilities:

```{r household_options, eval=FALSE}
# Using built-in Australian data
CQ.sim2(..., region = "NSW")
CQ.sim2(..., region = "Victoria")

# Using custom household distribution
uk_probs <- c(0.28, 0.35, 0.18, 0.12, 0.05, 0.015, 0.004, 0.001)
CQ.sim2(..., hh_probs = uk_probs)
```

## Interpreting Results

### Infection Potential in Quarantine (IPq)

IPq represents the expected number of tertiary infections from undetected secondary cases. It accounts for:
- How long infections remain undetected
- Generation interval distribution
- Transmission potential
- Vaccine effectiveness

**Lower IPq = more effective strategy**

### Mean Cases

The mean number of undetected secondary cases per index case. This complements IPq by showing the raw number of cases that escape detection.

## Next Steps

- See the **Model Details** vignette for in-depth explanation of the simulation model
- See the **Finding Optimal Strategies** vignette for systematic strategy comparison methods

## Quick Reference

```{r reference, eval=FALSE}
# Basic simulation (COVID-19 defaults)
sim <- CQ.sim2(n.ind, TP, k, p.vac.idx, p.vac.sc, vacc.cor,
               VE.trans, VE.inf, quarantine.duration,
               the.scenario, region)

# Simulation with custom disease parameters
sim <- CQ.sim2(n.ind, TP, k, p.vac.idx, p.vac.sc, vacc.cor,
               VE.trans, VE.inf, quarantine.duration,
               the.scenario,
               hh_probs = custom_probs,
               inc_meanlog = 0.34, inc_sdlog = 0.42,
               gi_meanlog = 0.91, gi_sdlog = 0.52)

# Evaluate testing
results <- CQ.sim.test.times(sim, n.ind, test.times, TP, VE.trans, the.scenario)

# Evaluate no testing
baseline <- CQ.sim.notest(sim, n.ind, TP, VE.trans, the.scenario)
```
